---
title: "passing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
library(ggthemes)
library(ggrepel)
```

```{r}
data <- read.csv(url('https://raw.githubusercontent.com/SportsInfoSolutions/AnalyticsChallenge2020/master/Data/AnalyticsChallenge2020Data.csv'))
passes <- data %>%
  filter(EventType == 'pass',
         Spike == 0)

```

```{r}
passes$EPA <- as.numeric(passes$EPA)
passes <- na.omit(passes)
passes$EPA <- as.numeric(passes$EPA)
```


```{r}
passes <- passes %>%
  mutate(
    GenSpot = case_when(
      TechniqueName == '0' | TechniqueName == '1' | TechniqueName == '2i'  ~ 'Middle',
      TechniqueName == '2' | TechniqueName == '3' | TechniqueName == '4i' | TechniqueName == '4' ~ 'Inside',
      TechniqueName == '5' | TechniqueName == '6' | TechniqueName == '7' | TechniqueName == '9' | TechniqueName == 'Outside' ~ 'Outside',
      TechniqueName == 'Off Ball' ~ 'Other'
    )
  )
```

```{r}
passes <- passes %>%
  mutate(
    PrsByMiddle = if_else(Pressure == 1 & GenSpot == 'Middle', 1, 0),
    PrsByInside = if_else(Pressure == 1 & GenSpot == 'Inside', 1, 0),
    PrsByOutside = if_else(Pressure == 1 & GenSpot == 'Outside', 1, 0),
    PrsByOff = if_else(Pressure == 1 & GenSpot == 'Other', 1, 0),
    NoPrs = if_else(PressureOnPlay == 0, 1, 0)
  )
```

```{r}
pass.epa.model <- lm(EPA ~ PrsByMiddle + PrsByInside + PrsByOutside + NoPrs , data = passes)
summary(pass.epa.model)
```

When the outside rusher is the one getting to the QB that results in the highest negative EPA impact. Probably because outside rushes are able to get to the QB quicker?

Why only use pressures? Your goal should be to generate pressure to affect QBs given what we've learned about the stbility of pressure rate opposed to more random variance in sacks(find a way to cite this).

```{r}
x_prs_lbls <- c('No', 'Yes')
prsInPlot <- passes %>%
  filter(PrsByInside == 1 | NoPrs == 1) %>%
  ggplot(aes(x=as.factor(PrsByInside), y=EPA, col=as.factor(PrsByInside))) +
  geom_boxplot(aes(fill=as.factor(PrsByInside)), alpha=.2) +
  theme_minimal() +
  scale_color_brewer(palette = 'Set2') +
  scale_fill_brewer(palette = 'Set2') +
  labs(title = 'Effect of Inside Pressure on Passing EPA', col='Pressure From Inside',
       x='Pressure From Inside', fill='Pressure From Inside') +
  theme(legend.position = 'none',
        plot.title = element_text(size = 16, hjust = 0.5, face = 'bold'),
        plot.subtitle = element_text(size = 14, hjust = 0.5)) +
  scale_x_discrete(labels= x_prs_lbls) +
  ylim(-3,3)
prsOutPlot <- passes %>%
  filter(PrsByOutside == 1 | NoPrs == 1) %>%
  ggplot(aes(x=as.factor(PrsByOutside), y=EPA, col=as.factor(PrsByOutside))) +
  geom_boxplot(aes(fill=as.factor(PrsByOutside)), alpha=.2) +
  theme_minimal() +
  scale_color_brewer(palette = 'Set2') +
  scale_fill_brewer(palette = 'Set2') +
  labs(title = 'Effect of Outside Pressure on Passing EPA', col='Pressure From Outside',
       x='Pressure From Outside', fill='Pressure From Outside') +
  theme(legend.position = 'none',
        plot.title = element_text(size = 16, hjust = 0.5, face = 'bold'),
        plot.subtitle = element_text(size = 14, hjust = 0.5)) +
  scale_x_discrete(labels= x_prs_lbls) +
  ylim(-3,3)
prsOutPlot
prsMidPlot <- passes %>%
  filter(PrsByMiddle == 1 | NoPrs == 1) %>%
  ggplot(aes(x=as.factor(PrsByMiddle), y=EPA, col=as.factor(PrsByMiddle))) +
  geom_boxplot(aes(fill=as.factor(PrsByMiddle)), alpha=.2) +
  theme_minimal() +
  scale_color_brewer(palette = 'Set2') +
  scale_fill_brewer(palette = 'Set2') +
  labs(title = 'Effect of Middle Pressure on Passing EPA', col='Pressure From Middle',
       x='Pressure From Middle', fill='Pressure From Middle') +
  theme(legend.position = 'none',
        plot.title = element_text(size = 16, hjust = 0.5, face = 'bold'),
        plot.subtitle = element_text(size = 14, hjust = 0.5)) +
  scale_x_discrete(labels= x_prs_lbls) +
  ylim(-3,3)
prsMidPlot
```

```{r}
most_prs <- passes %>%
  group_by(PrsByOutside, PrsByInside, PrsByMiddle) %>%
  summarise(Average_EPA = mean(EPA),
            SD_EPA = sd(EPA))
most_prs <- reshape2::melt(most_prs, id.vars = c('Average_EPA', 'SD_EPA'))
most_prs <- most_prs %>%
  filter(value==1)
prs_viz <- most_prs %>%
  ggplot(aes(
    x=variable, y=Average_EPA, col=variable, fill=variable
  )) +
  geom_col(alpha=.5) +
  scale_color_brewer(palette = 'Set1') +
  scale_fill_brewer(palette = 'Set1') +
  theme_minimal() +
  labs(title = 'Effect of Pressure on Passing EPA',
       x='Pressure Spot', y='Avg. Effect on EPA') +
  theme(legend.position = 'none',
        plot.title = element_text(size = 16, hjust = 0.5, face = 'bold'),
        plot.subtitle = element_text(size = 14, hjust = 0.5)) +
  scale_x_discrete(labels=c('Outside', 'Inside', 'Middle'))
ggsave('prsViz.png', prs_viz, dpi = 450)
```

```{r}
prs_spots <- cbind.data.frame(sum(passes$PrsByOutside), sum(passes$PrsByInside), sum(passes$PrsByMiddle))
colnames(prs_spots) <- list('Outside', 'Inside', 'Middle')
prs_count <- reshape2::melt(prs_spots) %>%
  ggplot(aes(
    x=variable, y=value, col=variable, fill=variable 
  )) +
  geom_col(alpha=.5) +
  scale_color_brewer(palette = 'Set1') +
  scale_fill_brewer(palette = 'Set1') +
  theme_minimal() +
  labs(title = 'Which Areas Generate the Most Pressures?',
       x='Pressure Spot', y='Amount of Pressures') +
  theme(legend.position = 'none',
        plot.title = element_text(size = 16, hjust = 0.5, face = 'bold'),
        plot.subtitle = element_text(size = 14, hjust = 0.5)) +
  scale_x_discrete(labels=c('Outside', 'Inside', 'Middle'))
prs_count
```

```{r}
prs_count_adj <- reshape2::melt(prs_spots)
sum(passes$GenSpot == 'Middle')
techs <- c(21383,12027,6911)
prs_count_adj <- cbind.data.frame(prs_count_adj, techs) %>%
  mutate(
    Adjusted_Amount = value / techs
  )
prs_count_adj <- reshape2::melt(prs_count_adj, id.vars=c('techs', 'variable'))
colnames(prs_count_adj) <- list('techs', 'variable', 'type', 'value')
prs_count_adj <- prs_count_adj[4:6,] %>%
  ggplot(aes(
    x=variable, y=value, col=variable, fill=variable 
  )) +
  geom_col(alpha=.5) +
  scale_color_brewer(palette = 'Set1') +
  scale_fill_brewer(palette = 'Set1') +
  theme_minimal() +
  labs(title = 'Average Pressure Rate by General Positioning',
       x='Pressure Spot', y='Pressure Rate') +
  theme(legend.position = 'none',
        plot.title = element_text(size = 16, hjust = 0.5, face = 'bold'),
        plot.subtitle = element_text(size = 14, hjust = 0.5)) +
  scale_x_discrete(labels=c('Outside', 'Inside', 'Middle')) #+
  #facet_grid(~ type)
prs_count_adj
ggsave('prsRate.png', prs_count_adj, dpi = 450)
```

Notice how on average, one outside technique will generate more pressurees than the middle

```{r}
pass.comp.model <- glm(as.numeric(Completion) ~ PrsByMiddle + PrsByInside + PrsByOutside + NoPrs, data = passes, family = 'binomial')
summary(pass.comp.model)
```

```{r}
PrsByMiddle <- c(1,0,0,0)
PrsByInside <- c(0,1,0,0)
PrsByOutside <- c(0,0,1,0)
NoPrs <- c(0,0,0,1)
pass_test_log <- cbind.data.frame(PrsByInside, PrsByMiddle, PrsByOutside, NoPrs)
pass_log_preds <- round(predict(pass.comp.model, pass_test_log, type = 'response'),4) * 100
pass_test_log <- cbind(pass_test_log, pass_log_preds)
gaps <- c('Middle', 'Inside', 'Outside', 'None')
gap_coords_x <- c(3,2,1, NA)
gap_coords_y <- c(1,1,1,NA)
pass_test_log <- cbind.data.frame(pass_test_log, gaps, gap_coords_x, gap_coords_y)
```

```{r}
pass_test_log[1:3,] %>%
  ggplot(aes(
    x=gap_coords_x, y=gap_coords_y, col=pass_log_preds, fill=pass_log_preds
  )) +
  theme_map() +
  geom_hline(yintercept = .495, col='blue', size=3) +
  geom_tile() +
  scale_color_gradient(low = 'forestgreen', high = 'firebrick') +
  scale_fill_gradient(low = 'forestgreen', high = 'firebrick')

```

```{r}
passes_outsideTechs <- passes %>%
  filter(
    TechniqueName == 5 | TechniqueName == 6 | TechniqueName == 7 | TechniqueName == 9 | TechniqueName == 'Outside'
  )
passes_outsideTechs <- passes_outsideTechs %>%
  mutate(
    PrsBy5 = if_else(Pressure == 1 & TechniqueName == 5, 1, 0),
    PrsBy6 = if_else(Pressure == 1 & TechniqueName == 6, 1, 0),
    PrsBy7 = if_else(Pressure == 1 & TechniqueName == 7, 1, 0),
    PrsBy9 = if_else(Pressure == 1 & TechniqueName == 9, 1, 0),
    PrsByOut = if_else(Pressure == 1 & TechniqueName == 'Outside', 1, 0)
  )
pass.epa.model.out <- lm(EPA ~ PrsBy5 + PrsBy6 + PrsBy7 + PrsBy9 + PrsByOut, data = passes_outsideTechs)
summary(pass.epa.model.out)
```

```{r}
pass.comp.model.out <- glm(as.numeric(Completion) ~ PrsBy5 + PrsBy6 + PrsBy7 + PrsBy9 + PrsByOut, data = passes_outsideTechs, family = 'binomial')
summary(pass.comp.model.out)
```


```{r}
pass.viz <- data.frame(pass.epa.model.out$coefficients)
pass.viz <- cbind(prsZone = rownames(pass.viz), pass.viz)
rownames(pass.viz) <- NULL
colnames(pass.viz) <- list('Pressure Zone', 'Effect on EPA')
pass.viz <- pass.viz %>%
  mutate(
    Position = case_when(
      `Pressure Zone` == 'PrsBy5' ~ '5-Tech',
      `Pressure Zone` == 'PrsBy6' ~ '6-Tech',
      `Pressure Zone` == 'PrsBy7' ~ '7-Tech',
      `Pressure Zone` == 'PrsBy9' ~ '9-Tech',
      `Pressure Zone` == 'PrsByOut' ~ 'Out'
    )
  )
pass_xs <- c(NA, 5,2.5,4.5,2,1)
pass_ys <- c(NA, 1,1,1,1,1)
oline_x <- c(4.5, 3.5, NA, NA, NA)
pass.viz <- cbind.data.frame(pass.viz, pass_xs, pass_ys)
```

```{r}
indViz <- pass.viz[2:6,] %>%
  ggplot(aes(
    x=pass_xs, y=pass_ys, col=`Effect on EPA`, shape=Position
  )) +
  theme_map() +
  geom_point(size=5) +
  geom_hline(yintercept = .95, col='blue', size=2, alpha=.5) +
  geom_hline(yintercept = .8, col='white') +
  geom_vline(xintercept = 6.2, col='white') +
  scale_color_gradient(low = 'green', high = 'red') +
  geom_text(x=5.2, y=.9, label='LT', col='black', size=9, fontface='bold') +
  geom_text(x=2.5, y=.9, label='TE', col='black', size=9, fontface='bold') +
  geom_text(x=5.865, y=.957, label='Line of Scrimmage', col='black', size=3) +
  geom_text_repel(aes(label = Position), size=7, point.padding = .5) +
  geom_segment(x=5, y=1, xend=6.15, yend=.85, col='red', arrow = arrow(length = unit(0.3, "inches"))) +
  geom_segment(x=4.5, y=1, xend=5.7, yend=.85, col='goldenrod3', arrow = arrow(length = unit(0.3, "inches"))) +
  geom_segment(x=2.5, y=1, xend=4.2, yend=.85, col='yellow4', arrow = arrow(length = unit(0.3, "inches"))) +
  geom_segment(x=2, y=1, xend=3.9, yend=.85, col='green', arrow = arrow(length = unit(0.3, "inches"))) +
  labs(title = "Effect of Each Outside Rusher",
       subtitle = 'Green = Best, Red = Worst') +
  theme(legend.position = 'none',
        plot.title = element_text(size = 16, hjust = 0.5, face = 'bold'),
        plot.subtitle = element_text(size = 14, hjust = 0.5))
indViz
ggsave('indViz.png', indViz, dpi = 450)  
?geom_text
```


make quiclk conditionally formatted tbl in excel to supplement

```{r}
# count(passes$TechniqueName)
# dis_techs <- ggplot(passes) +
#   geom_bar(aes(x=TechniqueName), col='blue', fill='blue', alpha=.5) +
#   theme_minimal() +
#   labs(title = 'Distribution of Techniques', y='Count', x='Technique') +
#   theme(legend.position = 'none',
#         plot.title = element_text(size = 16, hjust = 0.5, face = 'bold'),
#         plot.subtitle = element_text(size = 14, hjust = 0.5))
# 
# ggsave('techDis.png', dis_techs, dpi = 450)
```

```{r}
# dis_spots <- ggplot(passes) +
#   geom_bar(aes(x=GenSpot), col='forestgreen', fill='forestgreen', alpha=.5) +
#   theme_minimal() +
#   labs(title = 'Distribution of General Positions', y='Count', x='Technique') +
#   theme(legend.position = 'none',
#         plot.title = element_text(size = 16, hjust = 0.5, face = 'bold'),
#         plot.subtitle = element_text(size = 14, hjust = 0.5))
# ggsave('spotDis.png', dis_spots, dpi = 450)
```

